The Effects of Varying Depth in Cosmic Shear Surveys

### 2019-09-03 paperdraft.tex
added marika's paper

%
%-------------------------------------------------------------------
\section{Introduction}
The discovery of cosmic shear has provided us with a new and powerful cosmological tool to investigate the standard model of cosmology and to determine its parameters. Contrary to the analysis of the Cosmic Microwave Background (CMB) by \citet{2018arXiv180706209P}, cosmic shear is more sensitive to the properties of the local Universe and thus provides an excellent consistency check for the standard model. Current cosmic shear surveys are especially sensitive to the parameter $S_8=\sigma_8 \sqrt{\Omega_{\rm m}/0.3}$, where $\sigma_8$ denotes the normalization of the matter power spectrum and $\Omega_{\rm m}$ is the matter density.
It is interesting to note that all three current major cosmic shear results report a lower $S_8$ than inferred from the CMB analysis: While \citet{2018arXiv180706209P} determined a value of $S_8 = 0.830 \pm 0.013$, \citet{2018arXiv180909148H} report $S_8 = 0.800^{+0.029}_{-0.028}$ from analysis of the Subaru Hyper Suprime-Cam survey, \citet[][hereafter H18]{2018arXiv181206076H} obtained $0.737_{-0.036}^{+0.040}$ from KiDS+VIKING data and the Dark Energy Survey \citep[DES,][]{2018PhRvD..98d3528T} arrived at $S_8=0.782\pm 0.027$. Also, \citet{2013MNRAS.432.2433H} report $S_8 = 0.759 \pm 0.020$ from their analysis of CFHTLens data. Combined analyses of DES and KiDS data \citep{Joudaki:2019} yields even higher tension with the value determined for $S_8$ by the Planck Collaboration\todo{cite marikas new paper here as soon as it is there}. This discrepancy has received a lot of attention \citep{2017MNRAS.471.1259J,2016MNRAS.459..971K}. It could be interpreted as a statistical coincidence, a sign of new physics like massive neutrinos \citep{2014PhRvL.112e1303B}, time-varying dark energy or modified gravity \citep{2016A&A...594A..14P}; or as the manifestation of a systematic effect, either in the cosmic shear surveys or in the Planck mission \citep{2016ApJ...818..132A}, that is not yet accounted for.

@@

\begin{figure}
\centering
\includegraphics[width = \textwidth]{images/ebmodes0p5t72.pdf}
\caption{Difference in the E-modes (top left) and B-modes (bottom right) between the reference and the observed correlation functions. For comparison: Total E-modes of the reference correlation function $E_n$ (top left) and B-modes measured in the KV450 survey by Asgari et al. $B_n^{\rm{KV450}}$ (bottom right). All E- and B-modes were calculated using the logarithmic COSEBIS in S10 for an angular range of $\theta_{\rm{min}}=0.\!^\prime 5,\,\theta_{\rm{max}}=72\arcmin$.}
\label{fig:bmodes_cosebi}
\end{figure}

@@

We have shown that this effect can create B-modes. However, \todo{cite new asgari et al. here as soon as it is on arxiv}\citet{2018arXiv181002353A} measured the B-modes of the KV450 survey in the same $\theta$-range. Those B-modes are greater than the ones created by varying depth by a factor of approx.~$50$ and show a completely different pattern, so it is safe to say that the B-modes in the KV450 survey are not created by varying depth. An interesting observation is that the change in E-modes is as big as the created B-modes (compare Fig.~\ref{fig:bmodes_cosebi}). This means that as soon as this effect causes a significant bias in the cosmological parameters, it will also create significant B-modes\footnote{While this is no big surprise, it is not trivial. It could be possible that a systematic effect only creates E-modes and no B-modes, which would be extremely unfortunate as it could bias cosmological parameters without ever being detected by a B-mode analysis.}. Additionally, the created pattern is very characteristic, which makes it easy to recognize in a B-mode analysis of an actual survey.

>>>>

%
%-------------------------------------------------------------------
\section{Introduction}
The discovery of cosmic shear has provided us with a new and powerful cosmological tool to investigate the standard model of cosmology and to determine its parameters. Contrary to the analysis of the Cosmic Microwave Background (CMB) by \citet{2018arXiv180706209P}, cosmic shear is more sensitive to the properties of the local Universe and thus provides an excellent consistency check for the standard model. Current cosmic shear surveys are especially sensitive to the parameter $S_8=\sigma_8 \sqrt{\Omega_{\rm m}/0.3}$, where $\sigma_8$ denotes the normalization of the matter power spectrum and $\Omega_{\rm m}$ is the matter density.
It is interesting to note that all three current major cosmic shear results report a lower $S_8$ than inferred from the CMB analysis: While \citet{2018arXiv180706209P} determined a value of $S_8 = 0.830 \pm 0.013$, \citet{2018arXiv180909148H} report $S_8 = 0.800^{+0.029}_{-0.028}$ from analysis of the Subaru Hyper Suprime-Cam survey, \citet[][hereafter H18]{2018arXiv181206076H} obtained $0.737_{-0.036}^{+0.040}$ from KiDS+VIKING data and the Dark Energy Survey \citep[DES,][]{2018PhRvD..98d3528T} arrived at $S_8=0.782\pm 0.027$. Also, \citet{2013MNRAS.432.2433H} report $S_8 = 0.759 \pm 0.020$ from their analysis of CFHTLens data. Combined analyses of DES and KiDS data \citep{Joudaki:2019,2019A&A...624A.134A} yield even higher tension with the value determined for $S_8$ by the Planck Collaboration. This discrepancy has received a lot of attention \citep{2017MNRAS.471.1259J,2016MNRAS.459..971K}. It could be interpreted as a statistical coincidence, a sign of new physics like massive neutrinos \citep{2014PhRvL.112e1303B}, time-varying dark energy or modified gravity \citep{2016A&A...594A..14P}; or as the manifestation of a systematic effect, either in the cosmic shear surveys or in the Planck mission \citep{2016ApJ...818..132A}, that is not yet accounted for.

@@

\begin{figure}
\centering
\includegraphics[width = \textwidth]{images/ebmodes0p5t72.pdf}
\caption{Difference in the E-modes (top left) and B-modes (bottom right) between the reference and the observed correlation functions. For comparison: Total E-modes of the reference correlation function $E_n$ (top left) and B-modes measured in the KV450 survey by \citet{2019A&A...624A.134A} $B_n^{\rm{KV450}}$ (bottom right). All E- and B-modes were calculated using the logarithmic COSEBIS in S10 for an angular range of $\theta_{\rm{min}}=0.\!^\prime 5,\,\theta_{\rm{max}}=72\arcmin$.}
\label{fig:bmodes_cosebi}
\end{figure}

@@

We have shown that this effect can create B-modes. However, \citet{2019A&A...624A.134A} measured the B-modes of the KV450 survey in the same $\theta$-range. Those B-modes are greater than the ones created by varying depth by a factor of approx.~$50$ and show a completely different pattern, so it is safe to say that the B-modes in the KV450 survey are not created by varying depth. An interesting observation is that the change in E-modes is as big as the created B-modes (compare Fig.~\ref{fig:bmodes_cosebi}). This means that as soon as this effect causes a significant bias in the cosmological parameters, it will also create significant B-modes\footnote{While this is no big surprise, it is not trivial. It could be possible that a systematic effect only creates E-modes and no B-modes, which would be extremely unfortunate as it could bias cosmological parameters without ever being detected by a B-mode analysis.}. Additionally, the created pattern is very characteristic, which makes it easy to recognize in a B-mode analysis of an actual survey.
